---
title: "Homework 6"
author: "Ashwini Varghese"
date: "2022-12-03"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(mgcv)
library(modelr)
library(viridis)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.asp = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1



## Problem 2

We will start by loading the homicides data. 

```{r}
homicides = read.csv("./homicide-data.csv")
```

Next we will do some data cleaning as below:

```{r}
homicides =   
  homicides %>%
  unite("city_state", city:state, sep = ", ", remove = FALSE) %>% 
  mutate(solved = as.numeric(disposition == "Closed by arrest")) %>%
  filter(!(city %in% c("Dallas", "Phoenix", "Kansas City"))) %>%
  filter(city_state != "Tulsa, AL") %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  filter(victim_race %in% c("White", "Black")) %>%
  mutate(victim_race = fct_relevel(victim_race, "White")) %>%
  filter(victim_sex != "Unknown")

homicides %>% 
  group_by(city_state) %>%
  summarize() %>% 
  print(n = 51)

homicides %>% 
  group_by(victim_race) %>%
  summarise() 
  
```

Now for just the city of Baltimore, we will fit a logistic model with the homicide being solved as the outcome and with victim age, sex, and race as predictors. 

```{r}

Balt =
  homicides %>%
  filter(city == "Baltimore") %>%
  select(solved, victim_age, victim_race, victim_sex)

fit_balt = 
  Balt %>% 
  glm(solved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 

fit_balt %>% 
  broom::tidy() %>%
  mutate(lower_CI = exp(estimate - 1.96*std.error),
         upper_CI = exp(estimate + 1.96*std.error),
         OR = exp(estimate)) %>%
  select(term, OR, lower_CI, upper_CI) %>% 
  knitr::kable(digits = 3)


```

Let's repeat this for all the cities and get the adjusted OR with their 95% CI for solving homicides comparing males to females.  

```{r}
homi_nest =
  homicides %>% 
  select(city_state, solved, victim_age, victim_race, victim_sex) %>% 
  relocate(city_state) %>% 
  nest(data = solved:victim_sex)


fit_all = 
  homi_nest %>% 
  mutate(
    models = map(.x = data, ~glm(solved ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)) %>% 
  select(-data, -models) %>% 
  unnest(results) %>% 
  mutate(lower_CI = exp(estimate - 1.96*std.error),
         upper_CI = exp(estimate + 1.96*std.error),
         OR = exp(estimate)) %>%
  select(city_state, term, OR, lower_CI, upper_CI) %>%
  filter(term == "victim_sexMale")

fit_all %>%
  knitr::kable(digits = 3)


```





